---
title: "POL 150B/355B Homework 4"
date: "Due: 3/1/2018"
author: "Tongtong Zhang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = T,
	warning = F
)
```

In this homework, we'll be combining our knowledge of dictionary methods and discriminating words to create a custom scoring system of news articles.

The corpus contains articles about women in Middle Eastern and Western countries (not including the United States), published in the New York Times and Washington Post from 1980 to 2014. Our goal is to compare (and score) articles by the region they cover. In other words, how does U.S. news media portray women in the Middle East, compared to women in Western Europe and Canada?

We'll be working with two data frames:

1. `women` contains the training set. Relevant variables include:
- text: The full text of each article
- text.no.noun: The full text each article with proper nouns (names of people, places) removed.
- region: The geographic region the article is about, i.e. "MENA" (Middle East North Africa) or "West" (Canada and Western Europe.)

2. `unknown`: contains the "test set" of articles. It contains the same variables as `women`, but without the `region` column.

Load the following packages to being.

```{r}
library(tm)
require(matrixStats) # for statistics

rm(list=ls())
setwd("~/Dropbox/berkeley/Teaching/PoliSci_150B/Homeworks/HW_4")
```

# 1

## a. 

Load the `women` dataset into R. 

```{r}
d<-read.csv("women.csv",header=T)
```

## b. 

Create a document term matrix from the `text` column. The DTM should convert all words to lowercase; remove stop words, numbers and punctuation; and stem each term.

```{r}
docs <- Corpus(VectorSource(d$text)) #prepare a corpus
# create DTM with all the pre-processing requirements stated above
dtm <- DocumentTermMatrix(docs,
           control = list(stopwords = TRUE,
                          tolower = TRUE,
                          removeNumbers = TRUE,
                          removePunctuation = TRUE,
                          stemming=TRUE))
class(dtm)
```

## c. 

Using the `removeSparseTerms` function, remove sparse terms to result in .99 sparsity.

```{r}
ncol(dtm)
dim(dtm)

inspect(dtm[,1:5])

dtm <- removeSparseTerms(dtm,.99) #remove sparse terms that results in .99 sparsity or above
ncol(dtm) #only 591 columns remain
```

## d. 

Finally, convert the DTM to a data frame.

```{r}
dtm<-as.data.frame(as.matrix(dtm)) #convert matrix to dataframe
```

# 2. Find words distinctive to "MENA" and "West"

## a. 

In lecture, we discussed the "Standardized Mean Difference" score measuring word distinctiveness. Using this score, find the top 10 distinctive words for the MENA and West regions, respectively.

```{r}
#subset to MENA and West regions respectively
MENA<-dtm[which(d$region=="MENA"),]
west<-dtm[which(d$region=="West"),]

#For each word, create mean frequency in MENA and West corpra respectively
total_MENA<-sum(colSums(MENA)) #total number of words in each corpus
total_west<-sum(colSums(west))
mu_mena<-colSums(MENA)/total_MENA
mu_west<-colSums(west)/total_west

#For each word, create variance in frequency across documents in each corpus respectively
var_mena<-colVars(as.matrix(MENA))
var_west<-colVars(as.matrix(west))

#Calculate the weight for each word
num<-mu_west-mu_mena
denom<-sqrt(var_mena+var_west)
weight<-num/denom

#Get the top 10 distinctive words in each corpus
names(sort(weight,decreasing=T)[1:10]) # top 10 words with the most positive weights are distinctive for West
names(sort(-weight,decreasing=T)[1:10]) #last 10 words with the most negative weights are distinctive for MENA

```

## b. 
Create an data frame called `dict` that contains the top 30 distinctive words for the MENA and West regions and their scores. In other words, the dataframe should have dimensions 60 x 2.

```{r}
weight<-sort(weight,decreasing=T)
head(weight,30) #top 30 distinctive words for West
tail(weight,30) #top 30 distinctive words for MENA

score<-c(head(weight,30) ,tail(weight,30))
dict<-as.data.frame(score)
dict$words<-row.names(dict)
```

# 3. Classification

## a. 

Load & preprocess the `unknown` dataset in the same fashion as you did `women`, above.

```{r}
unknown<-read.csv("unknown.csv",header=T)
docs <- Corpus(VectorSource(unknown$text)) #prepare a corpus
# create DTM with all the pre-processing requirements stated above
dtm_test <- DocumentTermMatrix(docs,
           control = list(stopwords = TRUE,
                          tolower = TRUE,
                          removeNumbers = TRUE,
                          removePunctuation = TRUE,
                          stemming=TRUE))

dtm_test <- removeSparseTerms(dtm_test,.99) #remove sparse terms
dtm_test<-as.data.frame(as.matrix(dtm_test)) #convert dtm to a dataframe
```

## b.

Using the `dict` you created above, score each one of the `unknown` documents. Store the scores as a new column in the `unknown` dataframe.

```{r}
#get the index of the columns corresponding to the 60 distinctive words in dict
index<-sapply(dict$words, FUN=function(x) which(colnames(dtm_test)==x)) 
#Subset dtm_test to only the 60 columns corresponding to the 60 distinctive words
test<-dtm_test[,index]

#Calculate the scores for each test document
score<-as.matrix(test) %*% dict$score

#Attach the scores to the end of "unknown" dataframe
unknown$score<-as.numeric(score)
```

## c.

Create another column called `label` in `unknown` that classifies each article using it's score. Assign label = "MENA" if the score is less than 0, and "West" if the score is greater than 0. 

```{r}
unknown$label<-ifelse(unknown$score <0, "MENA","West") 
```

## d.

We've given you the correct labels below. How many did you get right using this method?

```{r}
true.labels <- c("MENA", "West", "MENA", "MENA", "MENA", "MENA", "West", "West", "MENA", "MENA", "MENA", "MENA", "MENA", "MENA", "West", "West", "West", "MENA", "MENA", "MENA", "West", "MENA", "West", "West", "West", "West", "West", "MENA", "West", "MENA", "MENA", "MENA", "MENA", "West", "MENA", "West", "West", "MENA", "MENA", "West", "MENA", "West", "West", "West", "West", "West", "West", "West", "MENA", "West", "West", "West", "MENA", "MENA", "MENA", "MENA", "MENA", "MENA", "MENA", "West", "West", "MENA", "West", "MENA", "West", "West", "West", "MENA", "West", "West", "MENA", "West", "West", "West", "West", "West", "MENA", "MENA", "MENA", "MENA", "MENA", "West", "West", "West", "MENA", "West", "West", "West", "West", "West", "MENA", "MENA", "West", "West", "MENA", "MENA", "MENA", "MENA", "MENA", "West")

#Calculate the number of accurate classifications => We got everything right!!!!
sum(true.labels==unknown$label)

```