---
title: "Interactive Code Lecture 10"
author: "150B/355B Introduction to Machine Learning"
date: "2/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Comparing Songs on the Thriller Album

To demonstrate sentiment analysis, we're going to explore lyrics from Michael Jackson's Thriller album. 

Road the code below to get started.

```{r}
rm(list=ls())
setwd('~/YOUR/PATH/HERE/10_Dictionary')
library(tm)

thriller <- read.csv("thriller.csv")
```

## 1.1 

First we must preprocess the corpus. Create a document-term matrix from the `Lyrics` column of the `thriller` data frame. Complete the following preprocessing steps:
- convert to lower
- remove stop words
- remove numbers
- remove punctuation.

**Think**: Why do you think stemming is inappropriate for this application?

```{r}
# preprocess and create DTM
docs <- Corpus(VectorSource(thriller$Lyrics))

dtm <- DocumentTermMatrix(docs,
           control = list(tolower = TRUE,
                          removeNumbers = TRUE,
                          removePunctuation = TRUE,
                          stopwords = TRUE
                         ))
# convert to matrix, and from there convert again to data frame
dtm <- as.data.frame(as.matrix(dtm))
```

## 2. Setting up the sentiment dictionary

## 2.1

We're going to use sentiment dictionaries from the `tidytext` package. Install and load the package.

```{r}
install.packages("tidytext") # only install a package once!
library(tidytext) # load it each session
```
 
## 2.2 

Using the `get_sentiments` function, load the "bing" dictionary and store it in an object called `sent`. 

```{r}
sent <- get_sentiments("bing")
head(sent)
```

## 2.3

Add a column to `sent` called `score`. This column should hold a "1" for positive words and "-1" for negative words.

```{r}
sent$score <- ifelse(sent$sentiment=="positive", 1, -1)
```

## 3. Scoring the Thriller album

## 3.1 

We're now ready to score each song. 

(**NB**: There are probably many ways to program a script that performs this task. If you can think of a more elegant way, go for it!)

First, we'll create a dataframe that holds all the words in our dtm along with their sentiment score.

```{r}
# get all the words in our dtm and put it in a dataframe
words = data.frame(word = colnames(dtm))
head(words)

# get their sentiment scores
words <- merge(words, sent, all.x = T)
head(words)

# replace NAs with 0s
words$score[is.na(words$score)] <- 0
head(words)
```

## 3.2

We can now use matrix algebra (!!) to multiply our dtm by the scoring vector. This will return to us a score for each document (i.e., song).

```{r}
# calculate documents scores with matrix algebra! 
scores <- as.matrix(dtm) %*% words$score

# put it in the original documents data frame
thriller$sentiment <- scores
```

Which song is happiest? Go listen to the song and see if you agree.

## 4. Making a function

## 4.1 

Using the code we wrote above, make a function that accepts 1) a vector of texts, and 2) a sentiment dictionary (i.e. a data frame with words and scores), and returns a vector of sentiment scores for each text

```{r}
sentiment_score <- function(texts, sent_dict){
  
  # YOUR CODE HERE
  
  return(scores)
  
}

# uncomment the line below to test it out!
# sentiment_score(thriller$Lyrics, sent)
```

## 4.2 

Using the function you wrote above, score the Thriller album with the "afinn" dictionary. Compare the scores across the two different dictionaries.

```{r}
# first load the dictionary
afinn <- get_sentiments("afinn")
head(afinn)

# then run the function
sentiment_score(thriller$Lyrics, afinn)
```

