---
title: "Interactive Code Lecture 5"
author: "150B/355B Introduction to Machine Learning"
date: "1/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setting up Iraq War Vote dataset

### 1.1

We're going to analyze the Senate's vote on the Iraq war authorization.We'll find this data set in the PSCL package, which is called `pscl` in R.

To get started, we need to install and load the library. Execute the following code to install pscl.

```{r}
rm(list=ls())
setwd('~/Dropbox/berkeley/Teaching/PoliSci_150B/Lectures/05_Logit')
# install.packages('pscl')
```

Now, load the pscl into your R environment.

```{r}
library(pscl)
```

To load the data set, execute the following code.

```{r}
data(iraqVote)
```

Checking the column names of the data set we have:
- y : The vote on the Iraq war authorization
- state.abb : Name of the senator's state
- name : senator's name
- rep : an indicator = TRUE if the senator is a Republican, FALSE if not a Republican
- state.name : the name of the senator's state
- gorevote: the share of the two party vote cast for Al Gore in the 2000 election

### 1.2

Let's examine the bivariate relationship between Republicans and their authorization vote. To do this, use the `table` function. The syntax is `table(VAR_ON_ROWS, VAR_ON_COLS)`. How many Republicans voted against the authorization?

```{r}
table(iraqVote$rep, iraqVote$y)
```

### 1.3

Using the data set, find the Republican who voted against the authorization.
```{r}
iraqVote[iraqVote$rep==TRUE & iraqVote$y == 0, ]
```

### 1.4

Subset the data to just Democrats and make a plot of the iraq authorization vote against the share of the two party vote for Gore. 

What do you notice about Democrats who come from states where Gore performed better? Where Gore performed worse?
```{r}
#  subset for democrats
dems <- subset(iraqVote, rep==FALSE)

# create a scatterplot
plot(y = dems$y, x = dems$gorevote)
```

## 2. Prediction using linear models

### 2.1 

Let's fit a predictive model using a linear probability model. Using a linear regression, regress the vote against rep and gorevote.
```{r}
fit <- lm(y ~ rep + gorevote, data = iraqVote)
summary(fit)
```

### 2.2

Using the regression, calculate predicted probabilities for each of the observations.  You can do this by creating an appropriate matrix or with a for loop.
```{r}
# create matrix of predictors
indep.matrix <- cbind(1, iraqVote$rep, iraqVote$gorevote)

# multiply by coefficients
pred_prob_lm<- c(indep.matrix%*%fit$coef) 

# we can also take the fitted values from the model
fitted_lm <- fit$fitted.values

# are they the same?
head(pred_prob_lm) #note: `head` returns first 5 items from an object
head(fitted_lm)
```

### 2.3

Summarize the predicted probabilities. What do you notice about the predicted probabilities? (in particular, the maximum and minimum values)
```{r}
summary(pred_prob_lm)
```

### 2.4

Write a function that takes the predicted probabilities and a threshold, and returns a classification decision. Use that function to classify each senator as Yay or Nay on the Iraq vote.
```{r}
# function inputting probability vector and threshold, returns class vector
class_func <- function(prob, threshold){
  class <- ifelse(prob>threshold, 1, 0)
  return(class)
}

# estimate class on predicted probabilities
class_lm <- class_func(pred_prob_lm, 0.5)

# take a peek
head(class_lm)
```

### 2.5

What proportion of senators do you classify as a likely Iraq Vote? What proportion of Democrats and Republicans?
```{r}
# proportion of all senators
summary(class_lm)

# proportion of dems
sum(class_lm[iraqVote$rep==0]) / sum(iraqVote$rep==0)

# proportion of rep
sum(class_lm[iraqVote$rep==1]) / sum(iraqVote$rep==1)
```

## 3. Exploring Logit functions

### 3.1

The natural logarithm function in R is `log`. If unfamiliar, try applying log to a few numbers. For example what does the following yield?
```{r}
log(exp(5)) 
```

### 3.2 

Complete the code below to write a function for the logit function.
```{r}
logit_func<- function(p){
  ### YOUR CODE HERE
  return(out)
}
```

### 3.3 

The code below plots the logit function from 0 to 1.  Where does the logit function equal 0?  
```{r}
# get a sequence of poitns
p <- seq(0,1,len=1000)

# plot those points and logit
plot(logit_func(p) ~ p, xlab = 'Probability', ylab = 'Logit(p)', type='l' , lwd = 3)

# add lines
abline(h = 0)
abline(v = 0.5)
```

### 3.4 

Now we're going to write a function for the inverse logit function. Complete the code below.

```{r}
# inverse logit
logit_inv_func<- function(a){
  # YOUR CODE HERE
  return(out)
}
```

Plot the logit function over the range -4 to 4. What do you notice is true about the rate of the change of the function at 0? What about the rate of change at -4 and 4? 

```{r}
# plot
a_seq <- seq(-4, 4, len = 1000)
linv_out <- logit_inv_func(a_seq)

plot(linv_out~a_seq, type='l', xlab='a', ylab = 'logit_inv(a)')
abline(v = 0)
abline(h = 0.5)
```

## 4. Fitting a Logistic Regression

Fitting a logistic regression is a generalization of fitting a linear regression. Let's work through an example of a logistic regression together.

We're fitting a simple logistic regression of the vote decision against Republican. To do this, we'll use the `glm` function.
```{r}
rep_reg<- glm(y~rep, family = binomial, data = iraqVote)
```

- `glm` stands for generalized linear model.
- `y` is the dependent variable here.
- `rep` is the independent variable.

We specify `family = binomial` to let glm know that we're interested in a logistic regression. And just like `lm`, data is how we specify the data we'll use.

We can check what is available in the object:
```{r}
names(rep_reg)
```

The model produces two predicted probabilities---one for Democrats and one for Republicans. We can extract those predicted probabilities with:
```{r}
rep_reg$fitted.values
```

### 4.1:

Fit a logistic regression of `y` on `rep` and `gorevote`. Store it in an object called `gore_reg`.
```{r}
gore_reg <- # YOUR CODE HERE
```

### 4.2

Retrive the predicted probabilities from the model above and store it in an object `logist_gore_preds`.
```{r}
logist_gore_preds <- # YOUR CODE HERE
```

### 4.3

The code below plots the predicted probabilities from the linear model and the logistic regression. Using this plot, how do the predicted probabilities of the two functions differ?
```{r}
plot(pred_prob_lm ~ c(logist_gore_preds), xlab = 'Logistic predictions', ylab = 'Linear predictions')
abline(h =1)
abline(v = 1)
```

### 4.4

Now, using the probabilities and the classification function and threshold from 2.4, classify the senators using the logistic regression function.
```{r}
class_gore_preds <- # YOUR CODE HERE
```

### 4.5

Using `table`, compare the classification from the linear model and the logistic regression. What do you notice?
```{r}
# YOUR CODE HERE
```

### 5. Model Evaluation 

### 5.1 

We are now ready to begin evaluating our model. We want to write three functions to calculate:
1) Accuracy
2) Precision
3) Recall

Functions 1 and 2 are written for you. Write 3 on your own.

```{r}
accuracy <- function(predicted, true){
  score <- (sum(predicted & true) + sum(!predicted & !true)) / length(true)
  return(score)
}

precision <- function(predicted, true){
  score <- sum(predicted & true) / sum(predicted)
  return(score)
}

# WRITE A FUNCTION FOR RECALL HERE
```

### 5.2

The code below uses the functions compare the accuracy, precision, and recall of the LM classifications and the logistic regression predicions. On the basis of these scores, can you make a strong argument for selecting either model?  
```{r}
cat('Linear Regression', '\n')
cat('Accuracy', '\n'); accuracy(class_lm, iraqVote$y)
cat('Precision', '\n'); precision(class_lm, iraqVote$y)
cat('Recall', '\n'); recall(class_lm, iraqVote$y)

cat('Logistic Regression', '\n')
cat('Accuracy', '\n'); accuracy(class_gore_preds, iraqVote$y)
cat('Precision', '\n'); precision(class_gore_preds, iraqVote$y)
cat('Recall', '\n'); recall(class_gore_preds, iraqVote$y)
```


### 5.3 

Finally, what happens as we vary the threshold on our classification? Let's focus on the predictions from the logistic regression. 

Using a for loop, assess how the precision, recall, and f scores varies as the threshold moves from 0 to 1.

```{r}
# create a vector of different threshold possibilities
thresh <- seq(0.0, 1, len = 1000)

# loop through and calculate classes based on these probabilities
store_f <- store_prec <- store_rec <- c()
for(z in 1:length(thresh)){
  temp_class<- class_func(logist_gore_preds, thresh[z])
  store_prec[z]<- precision(temp_class, iraqVote$y)
  store_rec[z]<- recall(temp_class, iraqVote$y)
  store_f[z]<- store_prec[z]*store_rec[z]/(store_prec[z] + store_rec[z])
}
```

### 5.4 

What is the threshold that maximizes f? What do you notice as we trade off precision and recall?

```{r}
# YOUR CODE HERE
```

