---
title: "POL 150B/350B Homework 1"
date: "Due: 1/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
```

This assignment will analyze vote returns for California House elections and vote choice in a presidential election.  

# 2006 California Congressional Election Results

Our goal in this exercise is to calcuate the probability that a Democratic candidate wins a House seat in a ``swing district": one where the support for Democratic and Republican candidates is about equal and the incumbent is a Democrat.  

## 1 

Load the data set `ca2006.csv` from your computer into R. `ca2006` a slightly modified version of the 2006 House election return data from the PSCL library.

The data set contains the following variables:

- `district`: California Congressional district 
- `prop_d`: proportion of votes for the Democratic candidate
- `dem_pres_2004`: proportion of two-party presidential vote for Democratic candidate in 2004 in Congressional district
- `dem_pres_2000`: proportion of two-party presidential vote for Democratic candidate in 2000 in Congressional district
- `dem_inc`: An indicator equal to 1 if the Democrat is the incumbent
- `contested`: An indicator equal to 1 if the election is contested
- `open`:  An indicator equal to 1 if the election is contested

```{r}
# YOUR CODE HERE
```

## 2

Create a plot of the proportion of votes for the Democratic candidate, against the proportion of the two-party vote for the Democratic presidential candidate in 2004 (John Kerry) in the district. Be sure to clearly label the axes and provide an informative title for the plot

```{r}
# YOUR CODE HERE
```

## 3

Regress the proportion of votes for the Democratic candidate, against the proportion of the two-party vote for the Democratic presidential candidate in 2004 in the district. Print the results and add the bivariate regression to the plot.

```{r}
# YOUR CODE HERE
```

## 4

Using the bivariate regression and a function you have written yourself (not `predict`!), report the expected vote share for the Democratic candidate if dem pres 2004 = 0.5.

```{r}
# YOUR CODE HERE
```

## 5

Now, regress `prop_d` against: `dem_pres_2004`, `dem_pres_2000` and `dem_inc`.

```{r}
# YOUR CODE HERE
```

## 6

Using the multivariate regression from 5 and a function you have written yourself, report the expected vote share for the Democratic candidate if:

- `dem_pres_2004` = 0.5 
- `dem_pres_2000` = 0.5 
- `dem_inc` = 1

```{r}
# YOUR CODE HERE
```

## 7

We are often interested in characterizing the uncertainty in our estimates. One technique for estimating uncertainty is called the bootstrap. Here, we will walk through the steps to produce a function that will output a bootstrap estimate.

Do the following 1000 times (in a for loop):

a) Using the `sample` function, randomly select 53 rows -- the number of districts in California in 2006 -- with replacement.
b) Using the randomly selected data set, fit the bivariate and multivariate regressions.
c) Using the regression, predict the expected vote share for the Democratic candidate from both regressions, using the values and functions from 4) and 6)
d) Store the predictions from both regressions.

```{r}
# YOUR CODE HERE
```

## 8

Create histograms for both predictions.

```{r}
# YOUR CODE HERE
```

## 1.9

We will say the model predicts that the incumbent wins if the predicted vote share is greater than 50%. What proportion of time does each model above predict the incumbent will win?

```{r}
# YOUR CODE HERE
```