---
title: "POL 150B/350B Homework 1 Solutions"
author: Haemin Jee
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
```

This assignment will analyze vote returns for California House elections and vote choice in a presidential election.  

# 2006 California Congressional Election Results

Our goal in this exercise is to calcuate the probability that a Democratic candidate wins a House seat in a ``swing district": one where the support for Democratic and Republican candidates is about equal and the incumbent is a Democrat.  

## 1 

Load the data set `ca2006.csv` from your computer into R. `ca2006` a slightly modified version of the 2006 House election return data from the PSCL library.

The data set contains the following variables:

- `district`: California Congressional district 
- `prop_d`: proportion of votes for the Democratic candidate
- `dem_pres_2004`: proportion of two-party presidential vote for Democratic candidate in 2004 in Congressional district
- `dem_pres_2000`: proportion of two-party presidential vote for Democratic candidate in 2000 in Congressional district
- `dem_inc`: An indicator equal to 1 if the Democrat is the incumbent
- `contested`: An indicator equal to 1 if the election is contested
- `open`:  An indicator equal to 1 if the election is contested

```{r}
#### Libraries
library(foreign)
library(ggplot2)

# set working directory
setwd("/Users/haeminjee/Documents/Year 3 Winter Quarter/150B TA Folder/Homework Solutions/HW_1")

# load data
ca <- read.csv("ca2006.csv") 

# examine structure of the data
str(ca)

# first observations
head(ca)

# summary statistics of variables
summary(ca)

```

## 2

Create a plot of the proportion of votes for the Democratic candidate, against the proportion of the two-party vote for the Democratic presidential candidate in 2004 (John Kerry) in the district. Be sure to clearly label the axes and provide an informative title for the plot

```{r}
plot1 <- ggplot(ca, aes(x=dem_pres_2004, y = prop_d)) + geom_point(size = 2) + theme_bw() +
  labs(x= "% votes for Kerry, 2004", 
       y = "% votes Dem candidate, 2006", 
       title = "Democratic vote share by county 2004, 2006")
plot1
```

## 3

Regress the proportion of votes for the Democratic candidate, against the proportion of the two-party vote for the Democratic presidential candidate in 2004 in the district. Print the results and add the bivariate regression to the plot.

```{r}
# the bivariate regression
bivariate <- lm(prop_d~dem_pres_2004, data = ca) 
summary(bivariate)
```

The summary of the bivariate model reveals that percent of votes for Kerry in a county has a positive and statistically significant relationship with the 2006 percent of votes for the Democratic candidate in 2006.

```{r}
# adding the regression line to plot
plot2 <- plot1 + geom_smooth(method = "lm", se = TRUE)
plot2 
```

## 4

Using the bivariate regression and a function you have written yourself (not `predict`!), report the expected vote share for the Democratic candidate if dem pres 2004 = 0.5.

```{r}
predict_mine <- function(value, model){
  coef <- model$coefficients
  y <- unname(coef[1] + coef[2]*value)
  return (y)
}

### Arguments: 
### value is the independent variable value for the observation we want to make
### a prediction for
### Model: the name of the model used to generate the coefficients
### Return: returns a district's predicted values of the proportion of votes for the ### Democratic candidate in 2006 


# predicted is the predicted value for the proportion vote share of the 
# Congressional Democratic candidate in 2006 using the bivariate model 
# if the vote share for Kerry in 2004 was .5
predicted <- predict_mine(.5, bivariate) 
print (predicted)
```
If the vote share for Kerry in a county is 0.5, then the expected vote share for the Democratic candidate is 0.537.

## 5

Now, regress `prop_d` against: `dem_pres_2004`, `dem_pres_2000` and `dem_inc`.

```{r}
multivariate <- lm(prop_d ~ dem_pres_2004 + dem_pres_2000 + dem_inc, data = ca)
summary(multivariate)
```
After adding other variables to the regression model, the only variable that remains statistically signifiacnt is `dem_inc`. 

## 6

Using the multivariate regression from 1.5 and a function you have written yourself, report the expected vote share for the Democratic candidate if:

- `dem_pres_2004` = 0.5 
- `dem_pres_2000` = 0.5 
- `dem_inc` = 1

```{r}
multipredict_mine <- function(values, model){
  values <- c(1, values)
   y <- values%*%unname(model$coefficients)
   return (y)
}

### Arguments: values is a vector of independent variable values for the observation ### we want to make a prediction for
### Model: the name of the model used to generate the coefficients
### Return: returns a distrit's predicted values of the proportion of votes for the ### Democratic candidate in 2006

multi_prediction <- multipredict_mine(c(.5, .5, 1), multivariate)
print(multi_prediction)

```
The expected vote share for the Democratic candidate is now 0.615.

## 7

We are often interested in characterizing the uncertainty in our estimates. One technique for estimating uncertainty is called the bootstrap. Here, we will walk through the steps to produce a function that will output a bootstrap estimate.

Do the following 1000 times (in a for loop):

a) Using the `sample` function, randomly select 53 rows -- the number of districts in California in 2006 -- with replacement.
b) Using the randomly selected data set, fit the bivariate and multivariate regressions.
c) Using the regression, predict the expected vote share for the Democratic candidate from both regressions, using the values and functions from 4) and 6)
d) Store the predictions from both regressions.

```{r}
set.seed(pi) 
# setting the seed for replication purposes

n <- 1000

# creating empty vectors for results
bi_results <- rep(NA)
multi_results <- rep(NA)

for (i in 1:n){
  
  # sample from the data
  sim_data <- ca[sample(nrow(ca), 53, replace = TRUE), ]
  
  # run bivariate and multivariate regressions
  bi <- lm(prop_d ~ dem_pres_2004, data = sim_data)
  multi <- lm(prop_d ~dem_pres_2004 + dem_pres_2000 + dem_inc, data = sim_data)
  
  # use my functions to make predictions and store
  bi_results[i] <- predict_mine(.5, bi)
  multi_results[i] <- multipredict_mine(c(.5, .5, 1), multi)
}

# putting predictions in table for ggplot
predicted_vals <- as.data.frame(cbind(bi_results, multi_results))
```

## 8

Create histograms for both predictions.

```{r}
biplot <- ggplot(predicted_vals, aes(bi_results)) + geom_histogram(bins=50) +   theme_bw() + labs(x= "", 
       y = "", 
       title = "Predicted values from bivariate regressions")
biplot

multiplot <- ggplot(predicted_vals, aes(multi_results)) + geom_histogram(bins=50) +   theme_bw() + labs(x= "", 
       y = "", 
       title = "Predicted values from multivariate regressions")
multiplot
```

## 1.9

We will say the model predicts that the incumbent wins if the predicted vote share is greater than 50%. What proportion of time does each model above predict the incumbent will win?

```{r}
prop_win_bi <- sum(predicted_vals$bi_results>0.5) / nrow(predicted_vals)
prop_win_multi <- sum(predicted_vals$multi_results>0.5) / nrow(predicted_vals)

prop_win_bi
prop_win_multi
```
For the bivariate model, 98\% of the simulations predict that the Democratic candidate will win (for this particular county profile).
For the multivariate model, 100\% of the simulations predict that the Democratic candidate will win (for this particular county profile). 