---
title: "Interactive Code Lecture 4 / 5"
author: "150B/355B Introduction to Machine Learning"
date: "1/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setting up Iraq War Vote dataset

### 1.1

We're going to analyze the Senate's vote on the Iraq war authorization.We'll find this data set in the PSCL package, which is called `pscl` in R.

To get started, we need to install and load the library. Execute the following code to install pscl.

```{r}
rm(list=ls())
setwd('~/Dropbox/berkeley/Teaching/PoliSci_150B/Lectures/04_Classification')
install.packages('pscl')
```

Now, load the pscl into your R environment.

```{r}
library(pscl)
```

To load the data set, execute the following code.

```{r}
data(iraqVote)
```

Checking the column names of the data set we have:
- y : The vote on the Iraq war authorization
- state.abb : Name of the senator's state
- name : senator's name
- rep : an indicator = TRUE if the senator is a Republican, FALSE if not a Republican
- state.name : the name of the senator's state
- gorevote: the share of the two party vote cast for Al Gore in the 2000 election

### 1.2

Let's examine the bivariate relationship between Republicans and their authorization vote. To do this, use the `table` function. The syntax is `table(VAR_ON_ROWS, VAR_ON_COLS)`. How many Republicans voted against the authorization?

```{r}
# YOUR CODE HERE
```

### 1.3

Using the data set, find the Republican who voted against the authorization.
```{r}
# YOUR CODE HERE
```

### 1.4

Subset the data to just Democrats and make a plot of the iraq authorization vote against the share of the two party vote for Gore. 

What do you notice about Democrats who come from states where Gore performed better? Where Gore performed worse?
```{r}
# YOUR CODE HERE
```

## 2. Prediction using linear models

### 2.1 

Let's fit a linear probability model. Using a linear regression, regress the vote against `rep` and `gorevote`.

```{r}
# YOUR CODE HERE
```

### 2.2

Using the regression, calculate predicted probabilities for each of the observations.  You can do this by creating an appropriate matrix or with a for loop.

```{r}
# YOUR CODE HERE
```

### 2.3

Summarize the predicted probabilities. What do you notice about the predicted probabilities? (in particular, the maximum and minimum values)
```{r}
# YOUR CODE HERE
```

### 2.4

Write a function that takes the predicted probabilities and a threshold, and returns a classification decision. Use that function to classify each senator as Yay or Nay on the Iraq vote.
```{r}
# YOUR CODE HERE
```

### 2.5

What proportion of senators do you classify as a likely Iraq Vote? What proportion of Democrats and Republicans?
```{r}
# YOUR CODE HERE
```


