---
title: "Interactive Code Lecture 3"
author: "150B/355B Introduction to Machine Learning"
date: "1/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setting up the time for a change model

Download and create in object from the .csv file ``TimeChange.csv" into your computer. 

```{r}
rm(list=ls())
setwd('~/Dropbox/berkeley/Teaching/PoliSci_150B/Lectures/03_Multiple')
elect<- read.csv('TimeChange.csv')
```

## 2. Fitting Multivariate Regression

### 2.1 

As a reminder, here's the bivariate regression we fitted in the last class:

```{r}
incumb_net<- lm(IncumbentVoteShare~Incumbent_Net_Approval, data = elect)
```

Using `lm`, regress `IncumbentVoteShare` on `Q1_GDP_Growth`, `Q2_GDP_Growth`, `Incumbent_Net_Approval`, and `Incumbent_Party_Two_Terms`. Summarize the output from the model.

```{r}
# fit model
time_for_change<- lm(IncumbentVoteShare~ Q1_GDP_Growth  + Q2_GDP_Growth + Incumbent_Net_Approval + Incumbent_Party_Two_Terms , data = elect)

# summarize
summary(time_for_change)
```

### 2.2

Compare the relationship between `Incumbent_Net_Approval` and `IncumbentVoteShare` in the bivariate and multivariate case.

```{r}
summary(incumb_net) # .161
incumb_net$coefficients[2]
summary(time_for_change) # 0.09511
```

## 3. Multivariate Prediction

### 3.1

Write a prediction function using the output from the multivariate regression.

```{r}
multi_pred<- function(obj, indep){
  coefs<- obj$coefficients
  pred<- coefs[1] + coefs[2] *indep[1] + coefs[3]*indep[2] + coefs[4]*indep[3] + coefs[5]*indep[4]
  return(pred)
}
```

### 3.2 

Compare it to fitted values.

```{r}
multi_pred_values <- multi_pred(time_for_change, elect[c(2:5)])
multi_pred_values
time_for_change$fitted.values
```


## 4. Linear Algebra in R ###

### 4.1

Create three vectors in R.  Make two of the vectors of length 4 and one vector of length 5.
- vector 1 should equal (1, 2, 3, 4)
- vector 2 should equal (3, 4, 5, 6)
- vector 3 should equal (7, 8, 9, 10, 11)

```{r}
test_vec1<- c(1, 2, 3, 4)
test_vec2<- c(3, 4, 5, 6)
test_vec3<- c(7,8, 9, 10, 11)
```
### 4.2

Confirm the length of the vectors using the length function.

```{r}
length(test_vec1)
length(test_vec2)
length(test_vec3)
```

### 4.3

Now, we're going to use `%*%` to take their inner product. First, take the inner product of vector 1 and vector 2. Second, write the inner product of vector 1 and vector 3. What do you notice? 

```{r}
# test_vec1%*%test_vec2
# test_vec1%*%test_vec3
```

## 5. Vector based regression

### 5.1 

Write a function to predict from a multivariate regression using vectors.

```{r}
multi_pred_vec<- function(obj, indep){
  coefs<- obj$coefficients
  indep.mat <- as.matrix(cbind(1, indep))
  pred<- c(indep.mat%*%coefs)
  return(pred)
}
```

### 5.2

Compare the predicted values between you two predict functions (one with linear algebra, one without) and from the lm ("canned") model.

```{r}
time_for_change$fitted.values
elect.ind.vars <- elect[c(2:5)]
multi_pred_vec(time_for_change, elect.ind.vars)
```







